D-Hammer: Efficient Equational Reasoning for Labelled Dirac Notation

### 2025-01-29 main.tex
.

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Labelled Dirac notation is a formalism commonly used by physicists to
represent many-body quantum systems and computer scientists to assert
properties of quantum programs. Labelled Dirac notation is supported
by a rich equational theory for proving equality between expressions
of the language. These proofs are typically carried on pen-and-paper,
and can be exceedingly long and error-prone. We introduce D-Hammer,
the first tool to supported automated equational proof for labelled
Dirac notation. The salient features of D-Hammer are: an expressive,
higher-order, dependently-typed language for labelled Dirac notation;
an efficient algorithm for normalization (incorporating associativity
and commutativity); and an optimized \CC\ implementation. We evaluate
the implementation on representative examples from plain and labelled
Dirac notation. In the case of plain Dirac notation, we show that our
implementation significantly outperforms DiracDec (Xu et al, POPL'25).

>>>>

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Labelled Dirac notation is a formalism commonly used by physicists to
represent many-body quantum systems and computer scientists to assert
properties of quantum programs. Labelled Dirac notation is supported
by a rich equational theory for proving equality between expressions
of the language. These proofs are typically carried on pen-and-paper,
and can be exceedingly long and error-prone. We introduce D-Hammer,
the first tool to supported automated equational proof for labelled
Dirac notation. The salient features of D-Hammer are: an expressive,
higher-order, dependently-typed language for labelled Dirac notation;
an efficient algorithm for normalization; and an optimized
\CC\ implementation. We evaluate the implementation on representative
examples from plain and labelled Dirac notation. In the case of plain
Dirac notation, we show that our implementation significantly
outperforms DiracDec (Xu et al, POPL'25).
